A quadratic-time algorithm for the cardinality-constrained college application problem

### 2022-07-11 paper_alma/paper.tex
skeleton of paper_alma/

\newif\ifen\enfalse
% Language selector; must be on line 3 for compile.sh to work
% Change to false to export Korean

\documentclass[11pt]{article} %try amsproc, amsart
\usepackage{kotex}
\usepackage{geometry}       \geometry{a4paper, margin=2.5cm}  
\usepackage{setspace}       \setstretch{1.1}
\usepackage{hyperref}       \hyperbaseurl{}    
                            \urlstyle{same}
\usepackage[ruled,linesnumbered]{algorithm2e}   \SetKwInOut{KwParams}{Parameters}
                                                \newcommand{\lIfElse}[3]{\lIf{#1}{#2 \textbf{else}~#3}}
\usepackage{stackengine}    \newcommand\xrowht[2][0]{\addstackgap[.5\dimexpr#2\relax]{\vphantom{#1}}}
\usepackage{fancyhdr}
\usepackage{graphicx}

% Author block
\usepackage{authblk}        \providecommand{\keywords}[1]{\textbf{Keywords:\ } #1}
\ifen
    \newcommand{\thetitle}{The College Application Problem}
    \newcommand{\therunningtitle}{\thetitle}
    \newcommand{\thedate}{\today}
    \newcommand{\theauthor}{Max Kapur}
    \newcommand{\theotherauthor}{Sung-Pil Hong}
\else
    \newcommand{\thetitle}{대학 입학 지원 최적화 문제}
    \newcommand{\therunningtitle}{\thetitle}
    \newcommand{\thedate}{\today}
    \newcommand{\theauthor}{Max Kapur}
    \newcommand{\theotherauthor}{홍성필}
\fi

\title{\thetitle}
\author{\theauthor}
\date{\thedate}

\ifen
\affil{Seoul National University\\
Department of Industrial Engineering\\
Management Science/Optimization Lab\\
Correspondence: \url{maxkapur@gmail.com}\\
Advisor: \theotherauthor}
\else
\affil{서울대학교\\
산업공학과 \\
경영과학/최적화 연구실\\
이메일: \url{maxkapur@gmail.com}\\
지도 교수: \theotherauthor}
\fi

% Header and footer
\pagestyle{fancy}
\renewcommand{\headrulewidth}{0pt}
\fancyhf{}
\lhead{\theauthor}
\chead{\thedate}
\rhead{\thetitle}
\cfoot{\thepage}

@@

\begin{abstract}
This paper considers the maximization of the expected maximum value of a portfolio of random variables subject to a budget constraint. We refer to this as the optimal college application problem. When each variable's cost, or each college's application fee, is identical, we show that the optimal portfolios are nested in the budget constraint, yielding an exact polynomial-time algorithm. When colleges differ in their application fees, we show that the problem is NP-complete. We provide four algorithms for this more general setup: a branch-and-bound routine, a dynamic program that produces an exact solution in pseudopolynomial time, a different dynamic program that yields a fully polynomial-time approximation scheme, and a simulated-annealing heuristic. Numerical experiments demonstrate the algorithms' accuracy and efficiency.

\begin{keywords}
submodular maximization, knapsack problems, approximation algorithms
\end{keywords}

\ifen\else The full text of this paper is available in English at \url{https://github.com/maxkapur/CollegeApplication}.\fi
\end{abstract}

\renewcommand{\abstractname}{요약}
\begin{abstract}
본 논문은 다수의 확률 변수로 구성된 포트폴리오의 기대 최댓값을 예산 조건 아래서 최대화하는 문제를 고려한다. 이를 대학 입학 지원 최적화 문제라고 부른다. 각 확률 변수의 비용, 즉 각 대학의 지원 비용이 동일한 경우, 최적 포트폴리오는 예산 제약식으로 결정된 포함 사슬 관계 성질을 가짐을 보이고 이를 바탕으로 다항 시간 해법을 제시한다. 대학의 지원 비용이 서로 다른 경우, 문제가 NP-complete함을 증명한다. 일반적인 문제를 위해 4가지 해법을 제시하며, 분지한계 기반 해법, 의사 다항 시간에 정확한 해를 출력하는 동적 계획 해법, 다른 동적 계획 기반으로 완전 다항 시간 근사 해법(fully polynomial-time approximation scheme), 그리고 모의 담금질(simulated annealing)을 이용한 휴리스틱 해법을 포함한다.
% 첫째는 분지한계 기반 해법이다. 둘째는 의사 다항 시간에 정확한 해를 출력하는 동적 계획 해법이다. 마지막으로 다른 동적 계획 기반으로 완전 다항 시간 근사 해법(fully polynomial-time approximation scheme)을 도출한다.
수리적 실험을 통해 알고리즘의 정확도와 효율성을 보인다.

\ifen 본 논문은 \url{https://github.com/maxkapur/CollegeApplication}에서 전문을 한글로 공유한다. \fi
\end{abstract}

@@

\pagebreak
\ifen \section{Introduction}  \else \section{서론} \fi
\ifen This paper considers the following optimization problem:
\else 본 논문은 다음과 같은 최적화 문제를 고려한다.\fi
%\begin{definition}[Optimal college application portfolio] \label{generalproblemstatement}
\begin{align} \label{headlineproblem}
\begin{split}
\text{maximize}\quad & v(\mathcal{X}) =  \operatorname{E}\Bigl[\max\bigr\{t_0,
\max\{t_j Z_j : j \in \mathcal{X}\}\bigr\}\Bigr] \\
\text{subject to}\quad & \mathcal{X} \subseteq \mathcal{C}, ~~\sum_{j\in \mathcal{X}} g_j \leq H
\end{split}
\end{align}
%\end{definition}
\ifen Here $\mathcal{C} = \{ 1 \dots m\}$ is an index set; $H > 0$ is a budget parameter; for $j = 1 \dots m$, $g_j > 0$ is a cost parameter and $Z_j$ is a random, independent Bernoulli variable with probability $f_j$; and for $j = 0\dots m$, $t_j\geq 0$ is a utility parameter. 
\else 단,  $\mathcal{C} = \{ 1 \dots m\}$은 지표 집합이며 $H$는 예산을 나타내는 모수다. 각 $j = 1 \dots m$에 대해 $g_j > 0$는 비용 모수이며 $Z_j$는 확률 $f_j$를 가지는 서로 독립적인 Bernoulli 변수다. 각 $j = 0 \dots m$에 대해 $t_j\geq 0$는 효용 모수다. \fi

@@

\begin{figure}[h!] 
 \centering
 \includegraphics[width=0.95\textwidth]{./plots/h_v-example.pdf}
  \caption{  \label{planetsplot}
  \ifen Application limit $h$ versus the optimal portfolio valuation $v^* = v(\mathcal{X}_h)$ in a fictional market with $m=8$ schools. The concave shape suggests the result of Theorem \ref{concavityinh}. 
  \else $m=8$개의 학교로 구성된 가상 입학 시상에서, 각 지원 제한 $h$에 해당하는 최적 포트폴리오의 가치 $v^* = v(\mathcal{X}_h)$. 곡선의 오목성은 정리 \ref{concavityinh}의 결과를 시사한다.\fi}
\end{figure}

@@

\begin{figure}[h!] 
\centering
\includegraphics[width=0.85\textwidth]{./plots/samplemarket.pdf}
  \caption{\label{samplemarket}
 \ifen A typical randomly-generated instance with $m=64$ schools and its optimal application portfolio. The application costs $g_j$ were drawn uniformly from $\{5, \dots, 10\}$. The optimal portfolio was computed using Algorithm \ref{ellisDP1}.
 \else $m=64$개의 학교로 구성된 전형적인 무작위로 생성한 인스턴스와 해당 최적 포트폴리오. 지원 비용 $g_j$는 $\{5, \dots, 10\}$에서 균일한 확률로 선정했으며 기호의 넓이와 비례된다. 최적 포트폴리오는 알고리즘 \ref{ellisDP1}(으)로 계산했다. \fi}
\end{figure}

@@

\begin{figure}[h!] 
\centering
\includegraphics[width=0.95\textwidth]{./plots/accuracy_simulatedannealing.pdf}
  \caption{\label{experiment3results}
 \ifen (Experiment 3.) Optimality ratio achieved by simulated annealing (Algorithm \ref{ellissimann}) in markets of varying size, with parameters $N = 500$, $T = 1/4$, and $r = 1/16$. Optimal portfolios were computed using Algorithm \ref{ellisDP1}. 
 \else  (실험 3.) 크기가 다양한 시장에 적용한 모의 담금질 (알고리즘 \ref{ellissimann}) 해법이 달성하는 최적성 비율. 모의 담금질의 모수는 $N = 500, T = 1/4, r = 1/16$이며, 최적 포트폴리오는 알고리즘 \ref{ellisDP1}(으)로 계산했다. \fi}
\end{figure}

@@

\parskip 0em
\leftskip 2em
\parindent -2em
\ifen \else
전민희. 2015. ``{[대입 수시 전략]} 총 6번의 기회 $\cdots$ `상향·소신·안정' 분산 지원하라.'' 중앙일보, 8월 26일. \url{https://www.joongang.co.kr/article/18524069}.\fi

Acharya, Mohan S., Asfia Armaan, and Aneeta S. Antony. 2019. ``A Comparison of Regression Models for Prediction of Graduate Admissions.'' In \emph{Second International Conference on Computational Intelligence in Data Science.} \url{https://doi.org/10.1109/ICCIDS.2019.8862140}.

\ifen Jeon, Minhee. 2015. ``[College application strategy] Six chances total\dots divide applications across reach, target, and safety schools'' (in Korean). Jungang Ilbo, Aug. 26. \url{https://www.joongang.co.kr/article/18524069}.\fi

>>>>

\documentclass[11pt]{article} %try amsproc, amsart
\usepackage{geometry}       \geometry{a4paper, margin=2.5cm}
\usepackage{setspace}       \setstretch{1.1}
\usepackage{hyperref}       \hyperbaseurl{}    
                            \urlstyle{same}
\usepackage[ruled,linesnumbered]{algorithm2e}   \SetKwInOut{KwParams}{Parameters}
                                                \newcommand{\lIfElse}[3]{\lIf{#1}{#2 \textbf{else}~#3}}
\usepackage{stackengine}    \newcommand\xrowht[2][0]{\addstackgap[.5\dimexpr#2\relax]{\vphantom{#1}}}
\usepackage[style=numeric]{biblatex}
\addbibresource{refs.bib}

% Author block
\usepackage{authblk}

\title{A quadratic-time algorithm for the cardinality-constrained college application problem}
\author[1]{Max Kapur}
\author[2]{Sung-Pil Hong}
\affil[1,2]{Department of Industrial Engineering, Seoul National University}
\date{\today}

@@

\begin{abstract}
%\input{abstract.tex}
This paper considers a novel portfolio optimization problem called the college application problem. We show that the objective function is a nondecreasing submodular set function and provide quadratic-time solution algorithm. We discuss the properties of the optimal solution and their implications for student welfare.

\textbf{Keywords:} combinatorial optimization, submodular maximization, portfolio optimization

\vfill
Correspondence may be addressed to Max Kapur.
\begin{itemize}
%\item[] Correspondence may be addressed to Max Kapur.

\item[] Email: \url{maxkapur@gmail.com}

\item[] Address: 39-411, 1 Gwanak-ro, Gwanak-gu, Seoul 08826, Republic of Korea
\end{itemize}

\end{abstract}

@@

%\input{body.tex}

\section{Introduction}

Due to the widespread perception that career earnings depend on college admissions outcomes, the solution of \eqref{headlineproblem} holds monetary value. In the American college consulting industry, students pay private consultants an average of \$200 per hour for assistance in preparing application materials, estimating their admissions odds, and identifying target schools (Sklarow 2018). In Korea, an important revenue stream for admissions consulting firms such as Megastudy (\url{megastudy.net}) and Jinhak (\url{jinhak.com}) is ``mock application'' software that claims to use artificial intelligence to optimize the client's application strategy.

Most quantitative research on college admissions to date has been descriptive in nature. A standard benchmark for logistic regression techniques, for example, involves estimating students' admissions probabilities from information about their past academic performance (Acharya et al. 2019; Lim 2013). When professional admissions consultants advise their clients on where to apply to college, however, they but typically rely on qualitative heuristics, such as balancing applications between safety and reach schools, rather than optimization models (Peck 2021). (Admissions consultants refer to schools with low utility and high admissions probability as \emph{safety schools}, and those with high utility and low admissions probability as \emph{reach schools.})

In the present study, we formulate the student's college application decision as the following optimization problem:
\begin{align} \label{headlineproblem}
\begin{split}
\text{maximize}\quad & v(\mathcal{X}) =  \operatorname{E}\Bigl[\max\bigr\{t_0,
\max\{t_j Z_j : j \in \mathcal{X}\}\bigr\}\Bigr] \\
\text{subject to}\quad & \mathcal{X} \subseteq \mathcal{C}, ~~|\mathcal{X}| \leq H
\end{split}
\end{align}
Here $\mathcal{C} = \{ 1 \dots m\}$ is an index set; $h > 0$ is the number of students the student can apply to; for $j = 1 \dots m$, $g_j > 0$ is a cost parameter and $Z_j$ is a random, independent Bernoulli variable with probability $f_j$; and for $j = 0\dots m$, $t_j\geq 0$ is a utility parameter.

Consider an admissions market with $m$ colleges. The $j$th college is named $c_j$. Consider a single prospective student in this market, and let each $t_j$-value indicate the utility she associates with attending $c_j$, where her utility is $t_0$ if she does not attend college. Let $g_j$ denote the application fee for $c_j$ and $H$ the student's total budget to spend on application fees. Lastly, let $f_j$ denote the student's probability of being admitted to $c_j$ if she applies, so that $Z_j$ equals one if she is admitted and zero if not. It is appropriate to assume that the $Z_j$ are statistically independent as long as $f_j$ are probabilities estimated specifically for this student (as opposed to generic acceptance rates). Then the student's objective is to maximize the expected utility associated with the best school she is admitted to. Therefore, her optimal college application strategy is given by the solution $\mathcal{X}$ to the problem above, where $\mathcal{X}$ represents the set of schools to which she applies. 
We focus on the special case in which students are constrained in the \emph{number} of schools to which they can apply, and show that the optimal portfolios in this case are nested in the budget constraint, implying the optimality of the greedy algorithm that adds schools one at a time according to how much they increase the objective value.

\subsection{Literature review}

The objective function is a nondecreasing submodular function in the sense first described by Nemhauser and Wolsey (1979). They showed that the greedy algorithm is a good approximation; we show that the same algorithm is  exact.

Fu (2014) considered a problem that is reducible to ours: since application is at cost rather than a constraint, the objective function is separable in the cost and therefore must be optimal over portfolios of the given size. Thus, solving her problem reduces to solving $m$ instances of \eqref{headlineproblem}.

To the best of our knowledge, the first systematic study of the college application problem was undertaken by Kapur in his master's thesis, which considers \eqref{headlineproblem} as well as the more general case of a knapsack constraint. The present study extends Kapur's results with a further discussion of the application strategies recommended in the admissions consulting industry and how these compare to the optimal strategy.

\subsection{Outline}

This paper has five sections. In section \ref{sectionModel}, we derive a closed-form expression for the objective function of \eqref{headlineproblem} and show that it is a monotonic submodular function. In section \ref{sectionGreedy}, we prove that the optimal portfolios are nested in the budget $h$, which implies the validity of the greedy algorithm. Further refinement reduces the computation time to $O(hm)$. In section \ref{sectionDiscussion}, we examine alternative solution mechanisms used in the admissions consulting industry and their performance relative to the optimal solution. A brief conclusion follows in section \ref{sectionConclusion}.

\section{Model} \label{sectionModel}

\section{The greedy algorithm} \label{sectionGreedy}

\section{Discussion} \label{sectionDiscussion}

\section{Conclusion} \label{sectionConclusion}

\section{References}

%\printbibliography

\parskip 0em
\leftskip 2em
\parindent -2em
Acharya, Mohan S., Asfia Armaan, and Aneeta S. Antony. 2019. ``A Comparison of Regression Models for Prediction of Graduate Admissions.'' In \emph{Second International Conference on Computational Intelligence in Data Science.} \url{https://doi.org/10.1109/ICCIDS.2019.8862140}.

Jeon, Minhee. 2015. ``[College application strategy] Six chances total\dots divide applications across reach, target, and safety schools'' (in Korean). Jungang Ilbo, Aug. 26. \url{https://www.joongang.co.kr/article/18524069}.

\ifen \section{Introduction}  \else \section{서론} \fi
\ifen This paper considers the following optimization problem:
\else 본 논문은 다음과 같은 최적화 문제를 고려한다.\fi
%\begin{definition}[Optimal college application portfolio] \label{generalproblemstatement}
\begin{align} \label{headlineproblem}
\begin{split}
\text{maximize}\quad & v(\mathcal{X}) =  \operatorname{E}\Bigl[\max\bigr\{t_0,
\max\{t_j Z_j : j \in \mathcal{X}\}\bigr\}\Bigr] \\
\text{subject to}\quad & \mathcal{X} \subseteq \mathcal{C}, ~~\sum_{j\in \mathcal{X}} g_j \leq H
\end{split}
\end{align}
%\end{definition}
\ifen Here $\mathcal{C} = \{ 1 \dots m\}$ is an index set; $H > 0$ is a budget parameter; for $j = 1 \dots m$, $g_j > 0$ is a cost parameter and $Z_j$ is a random, independent Bernoulli variable with probability $f_j$; and for $j = 0\dots m$, $t_j\geq 0$ is a utility parameter. 
\else 단,  $\mathcal{C} = \{ 1 \dots m\}$은 지표 집합이며 $H$는 예산을 나타내는 모수다. 각 $j = 1 \dots m$에 대해 $g_j > 0$는 비용 모수이며 $Z_j$는 확률 $f_j$를 가지는 서로 독립적인 Bernoulli 변수다. 각 $j = 0 \dots m$에 대해 $t_j\geq 0$는 효용 모수다. \fi

@@

\begin{figure}[h!] 
 \centering
% \includegraphics[width=0.95\textwidth]{./plots/h_v-example.pdf}
  \caption{  \label{planetsplot}
  \ifen Application limit $h$ versus the optimal portfolio valuation $v^* = v(\mathcal{X}_h)$ in a fictional market with $m=8$ schools. The concave shape suggests the result of Theorem \ref{concavityinh}. 
  \else $m=8$개의 학교로 구성된 가상 입학 시상에서, 각 지원 제한 $h$에 해당하는 최적 포트폴리오의 가치 $v^* = v(\mathcal{X}_h)$. 곡선의 오목성은 정리 \ref{concavityinh}의 결과를 시사한다.\fi}
\end{figure}

@@

\begin{figure}[h!] 
\centering
%\includegraphics[width=0.85\textwidth]{./plots/samplemarket.pdf}
  \caption{\label{samplemarket}
 \ifen A typical randomly-generated instance with $m=64$ schools and its optimal application portfolio. The application costs $g_j$ were drawn uniformly from $\{5, \dots, 10\}$. The optimal portfolio was computed using Algorithm \ref{ellisDP1}.
 \else $m=64$개의 학교로 구성된 전형적인 무작위로 생성한 인스턴스와 해당 최적 포트폴리오. 지원 비용 $g_j$는 $\{5, \dots, 10\}$에서 균일한 확률로 선정했으며 기호의 넓이와 비례된다. 최적 포트폴리오는 알고리즘 \ref{ellisDP1}(으)로 계산했다. \fi}
\end{figure}

@@

\begin{figure}[h!] 
\centering
%\includegraphics[width=0.95\textwidth]{./plots/accuracy_simulatedannealing.pdf}
  \caption{\label{experiment3results}
 \ifen (Experiment 3.) Optimality ratio achieved by simulated annealing (Algorithm \ref{ellissimann}) in markets of varying size, with parameters $N = 500$, $T = 1/4$, and $r = 1/16$. Optimal portfolios were computed using Algorithm \ref{ellisDP1}. 
 \else  (실험 3.) 크기가 다양한 시장에 적용한 모의 담금질 (알고리즘 \ref{ellissimann}) 해법이 달성하는 최적성 비율. 모의 담금질의 모수는 $N = 500, T = 1/4, r = 1/16$이며, 최적 포트폴리오는 알고리즘 \ref{ellisDP1}(으)로 계산했다. \fi}
\end{figure}

@@
